---
title: "p8105_hw6_cl4469"
author: "Chen Liang"
date: "2023-11-29"
output: github_document
---

```{r}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

#### Download Dataset
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

#### Bootstrap Resampling
```{r, warning=FALSE}
boot_strap = weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models= map(strap, ~lm(tmax ~ tmin + prcp, data = .x)),
    r2 = map(models, broom::glance),
    results = map(models, broom::tidy)) |>
  select(.id,results,r2) |> 
  unnest(r2,results) |>
  select(.id, term, estimate, r.squared) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |>
  rename(
    beta_0 = `(Intercept)`,
    beta_1 = tmin,
    beta_2 = prcp) |>
  summarise(id=.id,
            r_squared = r.squared,
            log_beta12 = log(beta_1*beta_2))

boot_noNA = boot_strap |>
  na.omit()
```

#### Plot the distribution of r_squared estimates
```{r}
boot_strap |>
  ggplot(aes(x = r_squared)) +
  geom_density()+
  labs(title = "Distribution of r_squared Estimates",
       x = "r_squared estimates")
```
Comment: We find that the distribution of r_squared estimates is approximately as normal distribution, and the r_squared estimates are around 0.92, indicating `tmin` and `prcp` are good predictors of `tmax`.

#### Plot the distributions of log(beta_1 * beta_2)
```{r}
boot_noNA |>
  ggplot(aes(x = log_beta12)) +
  geom_density() +
  labs(title = "Distribution of log(beta_1 * beta_2) Estimates",
       x = "log(beta_1 * beta_2) estimates")
```
Comment: After dropping the NA value of log_beta12 in boot_strap dataset, we plot the distribution of log(beta_1 * beta_2). We find that the estimates for log(beta_1 * beta_2) are left skewed and have some outliers, with a frequency around -5.5.

#### Calculate CI for r-squared and log_beta12
The 95% confidence interval for r_squared
```{r}
r_squared_ci = 
  quantile(pull(boot_strap, r_squared),
           probs = c(0.025,0.975)) |>
  knitr::kable()

r_squared_ci
```

The 95% confidence interval for log(beta_1 * beta_2)
```{r}
log_beta12_ci = 
  quantile(pull(boot_noNA, log_beta12),
           probs = c(0.025,0.975)) |> 
  knitr::kable()

log_beta12_ci
```

### Problem 3

#### Load and Clean birthweight dataset
```{r}
#load dataset
birth <- read_csv("data/birthweight.csv")

# Check for missing data
sum(is.na(birth))

# Clean dataset
birth_tidy = birth |>
  janitor::clean_names() |>
  mutate(babysex = case_when(babysex == 1 ~ 'male',
                             babysex == 2 ~ 'female'),
         frace = case_when(frace == 1 ~ "White",
                        frace == 2 ~ "Black",
                        frace == 3 ~ "Asian",
                        frace == 4 ~ "Puerto Rican",
                        frace == 8 ~ "Other",
                        frace == 9 ~ "Unknown"),
         malform = case_when(malform == 0 ~ 'absent',
                             malform == 1 ~ 'present'),
         mrace = case_when(mrace == 1 ~ "White",
                        mrace == 2 ~ "Black",
                        mrace == 3 ~ "Asian",
                        mrace == 4 ~ "Puerto Rican",
                        mrace == 8 ~ "Other"),
         babysex = as.factor(babysex),
         babysex = as.factor(babysex),
         malform = as.factor(malform),
         mrace = as.factor(mrace))
```
Comment: There's no NAs in the original dataset.

#### Regression Model
Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. 
```{r}
# fit a full model
full_model = lm(bwt ~ ., data = birth_tidy) |>
  step(direction = "backward", trace = FALSE) 

full_model |>
  broom::tidy() |>
  knitr::kable()
```
Comment: To creat a regression model for birthweight, I fit on a model based on all variables as predictors, including babysex, bhead, blength, delwt, fincome, gaweeks, mheight, etc. Then, I use backward stepwise regression.

#### Plot of Model Residuals Against Fitted Values
```{r}
birth_tidy |>
    add_predictions(full_model) |>
    add_residuals(full_model) |>
    ggplot(aes(x = pred, y = resid)) +
    geom_point()  + 
    geom_smooth(method = "lm") + 
    labs(title = "Residuals Against Fitted Values", 
       x = "Fitted Values", 
       y = "Residuals")
```
Comment: Most points are clustered between 2000 and 4000 fitted values, and between -1000 and 1000 residuals.Besides, the slope of the fitted line is almost 0. Since there is no clear direction for this plot, we cannot analysis the relationship for model residuals against fitted values.

#### Creat two other models
Compare your model to two others:One using length at birth and gestational age as predictors, One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
# model 2
model2 = lm(bwt ~ blength + gaweeks, data = birth_tidy)
model2 |>
  broom::tidy() |> 
  knitr::kable()

# model 3
model3 = lm(bwt ~ bhead * blength * babysex, data = birth_tidy)
model3 |>
  broom::tidy() |> 
  knitr::kable()
```

#### Make Comparison of Cross-validated Prediction Error
```{r}
cv_df= 
    crossv_mc(birth_tidy, 100) |>
    mutate(train = map(train, as_tibble),
           test = map(test,as_tibble)) |>
   mutate(model1 = map(train, ~ full_model),
          model2 = map(train, ~ model2),
          model3 = map(train, ~ model3)) |>
    mutate(rmse_model1 = map2_dbl(model1, test, ~rmse(model = .x,data = .y)),
           rmse_model2 = map2_dbl(model2, test, ~rmse(model = .x,data = .y)),
           rmse_model3 = map2_dbl(model3, test, ~rmse(model = .x,data = .y))
    )

cv_df |>
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```
Comment: From the plots, we can conclude that the full model, which consider all predictors, has the smallest RMSE, indicating that the predictions in this model are closer to the actual values than the other two; however, the model2, which consider length at birth and gestational age as predictors, has the largest RMSE, indicating that the predictions in this model may far away from the true values. Therefore, it is better to use model1 than model2 and model3.
